Name: Michael Labbe
Section: West
Date: 04/20/23
Assignment Name: Homework Thirteen
Goal of Program: Autocorrelation, AR/MA Models, ARIMA Modeling, Dickey-Fuller Unit Root Test and Ljung-Box Test.
Data Files Used: N/A

1.

```{r}
mean_y_t <- 0.5/(1-0.2)
mean_y_t
```
The mean of y_t is 0.625. 


2.
```{r}
var <- 4*(1 + 0.5^2)
var

#Calculate lag=0 autocorrelation
autocorrelation_0 <- 0.5/(1 + 0.5^2)
autocorrelation_0
```
  a. The mean of y_t is -2. 
  b. The variance of y_t is 5. 
  c. The lag=0 autocorrelation of y_t is 0.4.
  d. The lag-2 autocorrelation of y_t is 0.
  e. Yes, the time series y_t is stationary.


3.
A unit root test is a statistical test to determine if a time series is stationary, and if not, to determine the required order of differencing to obtain stationarity. In the augmented Dickey-Fuller test, the null hypothesis is that data is non-stationary and non-seasonal. If the null is rejected, then the time series is stationary. 


4.
The Ljung-Box test is useful in determining whether autocorrelations of several lags are zero. In a Ljung-Box Test, the null hypothesis is that all autocorrelations are zero. If the p-value <= 0.05, we have at least one non-zero serial correlation.