Name: Michael Labbe 
Date: 01/21/23 
Assignment: Homework One 
Goal of Program: Exploratory Data Analysis 
Data Files Used: faraway

```{r}
library(tidyverse)
#Load the data
install.packages("faraway", repos = "http://cran.us.r-project.org")
library("faraway")
data(uswages)
#Obtain numerical summaries
summary(uswages)
#Obtain number of rows
count(uswages)
#Obtain number of columns/variables
ncol(uswages)
#Look at scatterplot of educ vs wages
plot(uswages$educ,uswages$wage)
#Obtain correlation of education and wages
cor(uswages$educ,uswages$wage)
```

There are 2000 observations in the uswages data set.

There are 10 variables in the uswages data set.

Yes, the plot suggests an overall linear relationship between the education and wages variables with a couple of outliers.

There exists a weak, positive correlation between education and wages.

```{r}
#Fit simple linear model
par(mfrow=c(1,1))
plot(uswages$wage~uswages$educ)
mod1 <- lm(uswages$wage~uswages$educ)
summary(mod1)
abline(mod1)
```
Here, beta hat_0 (intercept) is 109.75 and the beta hat_1 (slope) is 38.01. This means that for x values of 0, we would expect y to be 109.75 on average and for every unit increase in x, we expect y to increase by 38.01 units on average. 


```{r}
#Compute the predicted value of wage for the first individual in the data set
y_hat <- 109.75 + 38.01 * uswages$educ[1]
y_hat

#Compute the residual
residual <- uswages$wage[1]-y_hat
residual

#Compute mean of all residuals
residuals <- uswages$wage - (109.75 + 38.01 * uswages$educ)
mean(residuals)
```
The predicted value of wage for the first individual in the data set is 793.93. 

The residual for the predicted value of 793.93 is -22.33. 

The mean of all the residuals is 0.0188. I think this value would vary slightly between different linear models depending on the strength of correlation between two variables. 

Definition of sampling distribution: If we sample many times from the population for a quantitative variable, where the population mean is mu and the population standard deviation is sigma, then the mean of all sample means (x-bars) is population mean mu. 

Applying sampling distribution concept to linear regression: If we sample many times from the population for a quantitative variable, where the population slope is beta_1, then the mean of all sample slopes (beta_1-hats) is population slope beta_1.


