Name: Michael Labbe
Section: West
Date: 04/27/23
Assignment Name: Homework Fourteen
Goal of Program: Transformations, Differencing, ACF/PACF plots, ARIMA vs. AUTO.ARIMA, Forecasting
Data Files Used: JnJ.csv

1.
```{r}
library(readr)
library(fpp)

setwd("C:\\Users\\malab\\OneDrive\\Documents\\Notre Dame\\Linear Models\\Data Sets")

jnj <- read_csv("JnJ.csv")
jnj <- dplyr::select(jnj,date,close)

jnj_ts <- ts(jnj$close)
plot.ts(jnj_ts)
```
From the time series plot, we can see a trend upwards over time for the close variable. 


2. Yes, a transformation is required to stabilize the variation.
```{r}
#The data shows different variation at different levels of the series, so a transformation is useful.
box_jnj <- BoxCox(jnj_ts,lambda=BoxCox.lambda(jnj_ts))
ts.plot(box_jnj)
```


3.
```{r}
adf.test(box_jnj) #p-value > 0.01: the time series is NOT stationary

ndiffs(box_jnj) #Check to see if we need regular differencing to obtain stationarity. Answer is YES

adf.test(diff(box_jnj)) #p-value < 0.01: the time series is now stationary
jnj_ts_diff <- diff(box_jnj)
ts.plot(jnj_ts_diff)
```


4.
```{r}
ggAcf(jnj_ts_diff)
ggPacf(jnj_ts_diff)
```
From the ACF and PACF plots, we can see that the suggested model is either an MA(2) model or an AR(2) model or some combination of both. 


5.
Two plausible models are c(2,1,0) and c(0,1,2) based on the plots in the previous answer. 
```{r}
#Try MA(2) model
fit <- Arima(box_jnj,order = c(0,1,2))
summary(fit)
#Try AR(2) model
fit2 <- Arima(box_jnj,order = c(2,1,0))
summary(fit2)
```
According to the AIC_c criterion the MA(2) model worked better than the AR(2) model. 


6.
```{r}
fit3 <- auto.arima(box_jnj, stepwise=FALSE)
summary(fit3) 
```
No, the auto.arima function chose a different model than both of the ones fit previously. In this particular example, the auto.arima function finds a model that would otherwise be difficult to find by just analyzing the PACF/ACF plots. 


7.
```{r}
#Check the residuals
tsdisplay(fit3$residuals) 
Box.test(fit3$residuals,lag=15, type='Ljung')
help(Box.test)
#p value is > 0.05, so residuals are white noise and the model fit is adequate.
```


8.
```{r}
#Plot the forecast
plot(forecast(fit3,h=10),include=80)
```